**Title:** Exploring the effects of BCG vaccination in patients diagnosed with tuberculosis: observational study using the Enhanced Tuberculosis Surveillance system

### 2018-03-29 output/Paper.Rmd
updated draft for grammar

BCG’s primary mode of action is to directly prevent the development of active, symptomatic disease. Its efficacy in adults is context specific, with estimates ranging between 0% and 78%.[@Mangtani2014a] Efficacy has been shown to be dependent on previous exposure, with unexposed individuals receiving the greatest benefit. Unlike in adults, BCG has consistently been shown to be highly protective against TB and TB meningitis in children.[@Rodrigues1993; @Colditz1994] For this reason the majority of countries that use BCG vaccinate at birth.[@Zwerling2011] Adult vaccination is no longer common, in the UK where universal BCG vaccination of adolescents was stopped in 2005 in favour of a targeted neonatal programme aimed at high risk children.

@@

This study used a large detailed dataset, with coverage across demographic groups, and standardized data collection from notifications and laboratories. The use of routine surveillance data means that this study would be readily repeatable with new data. The surveillance data contained multiple known risk factors, this allowed us to adjust for the confounding present in the univariable analysis, which attenuated the evidence for an association with BCG vaccination for all outcomes. However, there are important limitations to consider. The study was conducted within a population of active TB cases, therefore the association with all-cause mortality cannot be extrapolated to the general population. Additionally, vaccinated and unvaccinated populations may not be directly comparable because vaccination has been targeted at high-risk neonates in the UK since 2005. We mitigated this potential source for bias by conducting a sensitivity analysis including only those eligible for the universal school age scheme, and whilst the strength of associations were attenuated there remained some evidence of improved outcomes. Sensitivity analysis excluding recurrent cases indicated their inclusion may have biased our results towards the null.

@@

## Remove spurious results for deaths due to TB (in those who died) (no events in the data set to estimate on)
sum_imp_yrsinceBCG[2, c(2, 5)] <- '*Insufficient data*'

@@

```{r Imp ageatvac table, message=FALSE, echo=FALSE, results='asis', warning=FALSE}
  #BCG vaccination
if (imput) {
  sum_imp_ageatvac %>% 
              pretty_sum_results_table(footer = paste0('OR: odds ratio with 95% confidence intervals, aOR: adjusted odds ratio with 95% confidence intervals, fmi: fraction of missing information, * ', tb_death_footer),
                                       col_names =  c('Outcome', 'Age group (years)', 'OR (95% CI)', 'P-value', 'fmi', 'aOR (95% CI)', 'P-value', 'fmi'),
                                       label = 'sum_imp_ageatvac', 
                                       caption =   'Summary of associations between age at vaccination and all outcomes, using pooled imputed data(reference is vaccination at <1 year).', 
                                       cap_fun = SuptabRef)
}
```

>>>>

BCG’s primary mode of action is to directly prevent the development of active, symptomatic disease. Its efficacy in adults is context specific, with estimates ranging between 0% and 78%.[@Mangtani2014a] Efficacy has been shown to be dependent on previous exposure, with unexposed individuals receiving the greatest benefit. Unlike in adults, BCG has consistently been shown to be highly protective against TB and TB meningitis in children.[@Rodrigues1993; @Colditz1994] For this reason the majority of countries that use BCG vaccinate at birth.[@Zwerling2011] Adult vaccination is no longer common in the UK, where universal BCG vaccination of adolescents was stopped in 2005 in favour of a targeted neonatal programme aimed at high risk children.

@@

This study used a large detailed dataset, with coverage across demographic groups, and standardized data collection from notifications and laboratories. The use of routine surveillance data means that this study would be readily repeatable with new data. The surveillance data contained multiple known risk factors, this allowed us to adjust for these confounders in the multivariable analysis, which attenuated the evidence for an association with BCG vaccination for all outcomes. However, there are important limitations to consider. The study was conducted within a population of active TB cases, therefore the association with all-cause mortality cannot be extrapolated to the general population. Additionally, vaccinated and unvaccinated populations may not be directly comparable because vaccination has been targeted at high-risk neonates in the UK since 2005. We mitigated this potential source for bias by conducting a sensitivity analysis including only those eligible for the universal school age scheme, and whilst the strength of associations were attenuated there remained some evidence of improved outcomes. Sensitivity analysis excluding recurrent cases indicated their inclusion may have biased our results towards the null.

@@

## Remove spurious results for deaths due to TB (in those who died) (no events in the data set to estimate on)
sum_imp_yrsinceBCG[2, c(2, 5)] <- '*Insufficient data*'
sum_imp_yrsinceBCG[2, c(3, 6)] <- '-'

@@

```{r Imp ageatvac table, message=FALSE, echo=FALSE, results='asis', warning=FALSE}
  #BCG vaccination
if (imput) {
  sum_imp_ageatvac %>% 
              pretty_sum_results_table(footer = paste0('OR: odds ratio with 95% confidence intervals, aOR: adjusted odds ratio with 95% confidence intervals, fmi: fraction of missing information, * ', tb_death_footer),
                                       col_names =  c('Outcome', 'Age group (years)', 'OR (95% CI)', 'P-value', 'fmi', 'aOR (95% CI)', 'P-value', 'fmi'),
                                       label = 'sum_imp_ageatvac', 
                                       caption =   'Summary of associations between age at vaccination and all outcomes, using pooled imputed data (reference is vaccination at <1 year).', 
                                       cap_fun = SuptabRef)
}
```
